---
title: "Writeup Coursera"
author: "Marzena Magda"
date: "Sunday, December 21, 2014"
output: html_document
---

The project is related to activity recognition research - Qualitative Activity Recognition of Weight Lifting Exercises.
The objective of the analysis is to  predict how the activities were performed (if the activity was exactly according to the specifcation or with some mistakes)

The data set was uploaded from the site http://groupware.les.inf.puc-rio.br/har.

```{r}
training<-read.csv('C:\\Users\\Marzena\\Downloads\\pml-training.csv', header = TRUE, sep = ",", dec=",", stringsAsFactors=FALSE)
testing<-read.csv('C:\\Users\\Marzena\\Downloads\\pml-testing.csv', header = TRUE, sep = ",", dec=",", stringsAsFactors=FALSE)

```
The packages used: 
```{r}
library(caret)
library(randomForest)
```

First we did explanatory analysis. 
I performed data cleaning. 
I excluded variables with more than 50% of missing values.
The zero covariates analysis showed that we should remove the variables that could have no impact on the resluts such as new window and variables with many blanks. I also decided to exclude variables such as index, timestamps and user name because they can have no impact on activities so finally I have only numeric data.

```{r}
summary(training)
dim(training)
training1<-training[,colSums(is.na(training)) < 80]
nsv<-nearZeroVar(training1, saveMetrics=T)
numeric<-sapply(training1, is.numeric)
training2<-training1[,numeric]
training2<-training2[,4:29]
training3<-cbind(training$classe, training2)
names(training3)[1]<-"classe"

```
In order to predict the manner how the activities were done. I chose the random forest algorithm with classe as independent variable and the numeric variables we kept as predictors. I checked the importance of the predictors as well. The most important predictores were: num_window and magnet_dumbbell_y.
```{r}
rf1<-randomForest(classe~., data=training3)
rf1$importance
print(rf1)
```

Out of bag estimate of  error rate of the model is 0.37%. The accuracy table with classification error of each class is:

     A    B    C    D    E  class.error
A 5577    2    0    0    1 0.0005376344
B    1 3795    1    0    0 0.0005267316
C    0   19 3403    0    0 0.0055523086
D    1    0   37 3174    4 0.0130597015
E    0    1    1    5 3600 0.0019406709

To make predictions on testing data I used predict function:
```{r}
predictions<-predict(rf1,testing)
```


Source: Ugulino, W.; Cardador, D.; Vega, K.; Velloso, E.; Milidiu, R.; Fuks, H. Wearable Computing: Accelerometers' Data Classification of Body Postures and Movements. Proceedings of 21st Brazilian Symposium on Artificial Intelligence. Advances in Artificial Intelligence - SBIA 2012. In: Lecture Notes in Computer Science. , pp. 52-61. Curitiba, PR: Springer Berlin / Heidelberg, 2012. ISBN 978-3-642-34458-9. DOI: 10.1007/978-3-642-34459-6_6. 

Read more: http://groupware.les.inf.puc-rio.br/har#ixzz3MYuktyAW
